### 2022-02-14 misc/solutions/prob-method-sols.tex
Exercise 4.1

@@

>>>>

\def\given{\,|\,}

@@

\section 4. The Second Moment

\proclaim Exercise 4.1. Let $X$ be a random variable taking integral nonnegative values, let $\ex\{X^2\}$
denote the expectation of its square, and let $\var\{X\}$ denote its variance. Prove that
$$\pr\{ X = 0\} \le {\var\{X\} \over \ex\{X^2\}}.$$

\proof Since $X$ is integer and nonnegative, we have $\pr\{X = 0\} = 1-\pr\{X\ge 1\}$ and since
$\var\{X\} = \ex\{X^2\} - \ex\{X\}^2$, to get our result it suffices to show that
$$\pr\{X\ge 1\} \ge {\ex\{X\}^2 \over \ex\{X^2\}}.$$
We start by noting that
$$\ex\{X\} = \sum_{k=0}^\infty k \pr\{X=k\}
= \pr\{X\ge 1\}\sum_{k=1}^\infty {k\pr\{X=k\}\over \pr\{X\ge 1\}}
= \pr\{X\ge 1\}\ex\{ X \given X\ge 1\}.$$
Since the function $x\mapsto x^2$ is convex, we have, by Jensen's inequality,
$$\ex\{X\}^2 = \pr\{X\ge 1\}^2 \ex\{X\given X\ge 1\}^2 \le \pr\{X\ge 1\}^2 \ex\{X^2 \given X\ge 1\}
= \pr\{X\ge 1\} \ex\{X^2\}.$$
Dividing both sides by $\ex\{X^2\}$ gives us what we want.\slug
